Instant Continuation Marks--Just Add $\lambda$

### 2012-10-04 paper.tex
Added todo.
Good changes.

% motivate continuation marks
% motivate continuation marks in more languages
% discuss why they are not in more languages
% discuss approaches to getting them in languages with pros/cons
% ...culminating in our approach in terms of lambda

@@

We begin by further explaining continuation marks. We then formalize them via the introduction of a core and enhanced language with Redex models for each. We discuss the necessary properties for a meaningful transform.  We present two transformations of continuation marks--one direct and one CPS-like--to the call-by-value \lc. We then prove a meaning-preservation theorem.

\section{Continuation Marks}

@@

\section{Language Core and Extension}

@@

\section{Transformation}

@@

Our choice to keep the core language small by omitting lists as primitive values has the consequence of complicating our transform somewhat. Because lists are defined in terms of \lc\ values which are themselves touched by the transform and because of the commutativity property that $\mathcal{C}$ must satisfy, we cannot deal with a list of continuation marks directly--we must instead deal with a transformed list of transformed continuation marks.

Additionally, [values are often truncated, etc.]

\begin{equation}
\Cp{\abs{x}{e}}=\abs{x}{\C{e}}
\end{equation}

and we adjust $\hat{\mathcal{C}}$ so that
\begin{equation}
\Ch{p}=\app{\app{\C{p}}{\false}}{\Cp{\nil}}
\end{equation}
and recognize that a fully-realized transform must be made aware of this.

@@

%\subsection{Example: Factorial}

%For convenience, we will augment the base language with direct forms for expressing factorial. Define terms $e$ in \lv\ by
%\begin{equation}
%e=\app{e}{e}\,|\,(*\,e\,e)\,|\,\app{\mathrm{pred}}{e}\,|\,\app{\mathrm{zero?}}{e}\,|\,v\,|\,x
%\end{equation}
%and values $v$ by
%\begin{equation}
%v=\abs{x}{e}\,|\,\mathbb{N}
%\end{equation}
%.

%We can express the factorial function by the following
%\begin{schemedisplay}
%(Y (lambda (f)
%     (lambda (n)
%       (((zero? n)
%         1)
%        (* n (f (pred n)))))))
%\end{schemedisplay}

%and with continuation marks by
%\begin{schemedisplay}
%(Y (lambda (f)
%     (lambda (n)
%       (((zero? n)
%         (print (ccm) 1))
%        (* n (wcm n (f (pred n))))))))
%\end{schemedisplay}

@@

\section{Testing}

@@

\section{Proof}

@@

\section{Back to JavaScript}

@@

Factorial in this JavaScript extension can be implemented like the following:
\begin{verbatim}
var fac = function( n ) {
    if( n == 0 ) {
        console.log( ccm );
        return 1;
    }
    else {
        return wcm { n }
                   { n * fac( n - 1 ) };
    }
}
\end{verbatim}

A transformation of this function desugars these constructs into valid JavaScript.
\begin{verbatim}
var fac = function( flag, marks ) {
    return function( n ) {
        if( n == 0 ) {
            console.log( (function( flag, marks ) { return marks; })( false, marks ) );
            return 1;
        }
        else {
            return (function( flag, marks ) {
                return (function( mark_value, rest_marks ) {
                    return n * fac( false, [ mark_value ].concat( rest_marks ) )( n - 1 );
                })( n, f ? m.slice( 1 ) : m );
            })( flag, marks );
        }
    }
}
\end{verbatim}

>>>>

@@

We begin by further explaining continuation marks in section \ref{sec-continuation-marks}. We then formalize them via the introduction of a core and enhanced language in section \ref{sec-core-and-extension}. We discuss the necessary properties of a meaningful transformation between these languages. We present two transformations of continuation marks--one direct and one CPS-like--to the call-by-value \lc in section \ref{sec-transformation}. This is followed by a lightweight mechanized model for random testing in section \ref{sec-testing} and a proof of a meaning-preservation theorem in section \ref{sec-proof}. We finally apply it to JavaScript in section \ref{sec-javascript}.

\section{Continuation Marks}
\label{sec-continuation-marks}

@@

\section{Language Core and Extension}
\label{sec-core-and-extension}

@@

\section{Transformation}
\label{sec-transformation}

@@

Our choice to keep the core language small by omitting lists as primitive values has the consequence of complicating our transform somewhat. Because lists are defined in terms of \lc\ values which are themselves touched by the transform and because of the commutativity property that $\mathcal{C}$ must satisfy, we cannot deal with a list of continuation marks directly--we must instead deal with a transformed list of transformed continuation marks, and manipulation of this list within transformed terms must occur at the transformed level.

Additionally, after evaluation, values are ``truncated'' with their leading abstractions applied away. For instance, the transformation of the value \scheme|(lambda (x) x)| to \scheme|(lambda (flag) (lambda (marks) (lambda (x) (lambda (flag) (lambda (marks) x)))))| will yield, following evaluation, \scheme|(lambda (x) (lambda (flag) (lambda (marks) x)))|. For convenience, we define 
\begin{equation}
\Cp{\abs{x}{e}}=\abs{x}{\C{e}}
\end{equation}
and we adjust $\hat{\mathcal{C}}$ so that
\begin{equation}
\Ch{p}=\app{\app{\C{p}}{\false}}{\Cp{\nil}}
\end{equation}

@@

@@

% TODO
%explain transforms for values and variables
%fix references
%add example program
%fix introduction and handling of CPS
%talk about proof
%fix conclusion

\section{Testing}
\label{sec-testing}

@@

\section{Proof}
\label{sec-proof}

@@

\section{Back to JavaScript}
\label{sec-javascript}

@@

Tail-call elimination is not part of the JavaScript specification which complicates the treatment of the tail-call behavior. We have chosen to perform a simple static analysis which determines if the recursive call is in tail position. We simulate proper tail-call behavior by assigning the static value of the flag to reflect this.

A properly-recursive factorial can be implemented in this JavaScript extension as
\begin{verbatim}
var fac = function( n ) {
  if( n == 0 ) {
    console.log( ccm );
    return 1;
  }
  else {
    return wcm { n }
               { n * fac( n - 1 ) };
  }
}
\end{verbatim}

A transformation of this function desugars these constructs into vanilla JavaScript, leveraging native arrays for lists of continuation marks.
\begin{verbatim}
var fac = function( flag, marks ) {
  return function( n ) {
    if( n == 0 ) {
      console.log( (function( flag, marks ) { return marks; })( false, marks ) );
      return 1;
    }
    else {
      return (function( flag, marks ) {
        return (function( mark_value, rest_marks ) {
          return n * fac( false, [ mark_value ].concat( rest_marks ) )( n - 1 );
        })( n, flag ? marks.slice( 1 ) : marks );
      })( flag, marks );
    }
  }
}
\end{verbatim}

As expected, the output of a manual initiation of this function is
\begin{verbatim}
> fac( false, [] )( 5 )
[ 1, 2, 3, 4, 5 ]
120
\end{verbatim}

A tail-recursive factorial is expressible in this extension as
\begin{verbatim}
var fac = function( n, acc ) {
  if( n == 0 ) {
    console.log( ccm );
    return acc;
  }
  else {
    return wcm { n }
               { fac( n - 1, n * acc ) };
  }
}
\end{verbatim}
and transforms to
\begin{verbatim}
var fac = function( flag, marks ) {
  return function( n, acc ) {
    if( n == 0 ) {
      console.log( (function( flag, marks ) { return marks; })( false, marks ) );
      return acc;
    }
    else {
      return (function( flag, marks ) {
        return (function( mark_value, rest_marks ) {
          return fac( true, [ mark_value ].concat( rest_marks ) )( n - 1, n * acc );
        })( n, flag ? marks.slice( 1 ) : marks );
      })( flag, marks );
    }
  }
}
\end{verbatim}

The output of a manual initiation of this function is
\begin{verbatim}
> fac( false, [] )( 5, 1 )
[ 1 ]
120
\end{verbatim}
